### 2018-10-22 referee_respons.txt
Begin crafting referee response.

Thank you for the constructive criticisms and comments. We have significantly
revised the manuscript to improve its clarity and address several of your concerns.

Response to the specific concerns of the referee:

"A parton model of the proton has direct consequences not only for
heavy-ion collisions, but also for p+p collisions. The authors should
present predictions for p+p scattering using the most probable
parameters (or using the posterior distribution of parameters w, v,
n_c) for the differential elastic p+p scattering cross section. The
prediction can be compared to parametrized data at 5TeV. As a proxy
for the p+p data the parametrization from Phys. Rev. D88 (2013) 094019
can be used."

We agree that it would be interesting to explore the consequences of our
proton structure parameters on p-p observables. However, doing so properly
would involve significant work beyond the scope of the present study. We
therefore believe that such an analysis would be more appropriate for an
independent follow-up study. With this in mind, we've made all software, methods,
and data open source to the broader community.

"The posterior for n_c (the number of constituents) clearly shows
that the prior range is too constrained, a large part of the posterior
extends in the region n_c>9. I do not want to ask the authors to redo
the calculation with different prior, but a comment should be added in
text."

Thank you for pointing this out. We've added the following comment to
explain the truncation of the constituent number prior:

One exception is the constituent number nc which we limit for practical considerations.
Recall that each constituent fluctuates independently, weighted by a gamma random variable.
Hence for constituent numbers $n_c > 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish.
To counteract this effect, the constituent fluctuation variance must increase as nc increases.
Eventually, these required fluctuations become unreasonably large.
We find that for nc < 10, the energy density fluctuations are reasonable, and hence limit nc to this prior range."

"For the model predictions from the posterior experimental data (even preliminary) should be added, whenever available."

We've been careful to include all available data (public or otherwise), but it
is possible that we missed a dataset. In response to your suggestion, I've sent an email to the ALICE
collaboration to request their preliminary data, e.g. Pb-Pb mean pT, which could be
added to the posterior predictions where data is missing. Unfortunately, I have not heard back from them
yet, and I do not believe it is their policy to share preliminary data.
If you can direct me to a specific dataset, I would be happy to include it in
the maximum a posteriori observable plot.

Additional changes to the manuscript:

1. We've replaced the title with a more succinct version.

2. Several technical details of the initialization procedure have been
rewritten based on comments received from the community. The initial condition
parametrization is correspondingly recast as an initial energy density profile
(previously it was a parton density). We've also improved the
discussion of the initial free-streaming evolution and its matching to viscous
hydrodynamics.

3. We've simplified the nucleon substructure parameters, replacing the nucleon
width w [fm] and nucleon structure parameter X with a constituent sampling
radius r and constituent width v [fm]. The new parametrization is simpler and
less prone to misinterpretaiton. This transformation does not change the
conclusions of the study (nor should it).

4. We've corrected several inaccuracies that we made when discussing the
posterior nucleon width. In particular, the RMS radius differs from the Gaussian
width by a factor of sqrt(3).

5.