### 2019-04-29 paper/response.txt

Thank you for the constructive criticisms and comments. We have significantly revised the manuscript to improve its clarity and address several of your concerns.

Response to the specific concerns of the referee:

"A parton model of the proton has direct consequences not only for heavy-ion collisions, but also for p+p collisions. The authors should present predictions for p+p scattering using the most probable
parameters (or using the posterior distribution of parameters w, v, n_c) for the differential elastic p+p scattering cross section. The prediction can be compared to parametrized data at 5TeV. As a proxy
for the p+p data the parametrization from Phys. Rev. D88 (2013) 094019 can be used."

We agree that it would be interesting to explore the consequences of our proton structure parameters on p-p observables. However, doing so properly would involve significant work beyond the scope of the present study. We therefore believe that such an analysis would be more appropriate for an independent follow-up study. With this in mind, we've made all software, methods, and data open source to the broader community.

"The posterior for n_c (the number of constituents) clearly shows that the prior range is too constrained, a large part of the posterior extends in the region n_c>9. I do not want to ask the authors to redo the calculation with different prior, but a comment should be added in text."

Thank you for pointing this out. We've added the following comment to explain the truncation of the constituent number prior:

One exception is the constituent number $n_c$ which we limit for practical considerations. Recall that each constituent fluctuates independently, weighted by a gamma random variable. Hence for constituent numbers $n_c > 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish. To counteract this effect, the constituent fluctuation variance must increase as $n_c$ increases. Eventually, these required fluctuations become unreasonably large. We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence limit $n_c$ to this prior range.

"For the model predictions from the posterior experimental data (even preliminary) should be added, whenever available."

We've been careful to include all available data (public or otherwise), but it is possible that we missed a dataset. In response to your suggestion, we've sent an email to the ALICE collaboration to request their preliminary data, e.g. Pb-Pb mean pT, which could be added to the posterior predictions where data is missing. Unfortunately, I have not heard back from them yet, and I do not believe it is their policy to share preliminary data. If you can direct us to a specific dataset, we would be happy to include it in the maximum a posteriori observable plot.

Additional changes to the manuscript:

1. We've replaced the title with a more succinct version.

2. Several technical details of the initialization procedure have been rewritten based on comments received from the community. The initial condition parametrization is now presented as an energy density profile. Previously it was presented as a parton density with each parton carry some average energy per particle.  We've also improved the discussion of the initial free-streaming evolution and its matching to viscous hydrodynamics.

3. We've simplified the nucleon substructure parameters, replacing the nucleon width w [fm] and nucleon structure parameter X with a constituent sampling radius r and constituent width v [fm]. The new parametrization is simpler and less prone to misinterpretaiton. This transformation does not change the conclusions of the study (nor should it); we simply applied a transformation to the training data before emulation.

4. We've corrected several inaccuracies that were made when discussing the posterior nucleon width.